Between Subgraph Isomorphism and Maximum Common Subgraph

### 2016-09-02 paper/paper.tex
Wildcards as unit domains

@@

?? Note about the size of these instances, and how they're much larger than MCS instances.

@@

\begin{algorithm*}[tb]
\DontPrintSemicolon
\begin{multicols}{2}
\nl $\FuncSty{klessSubgraphIsomorphism}$ (\\\hspace*{2em}Graph $\mathcal{P}$, Graph $\mathcal{T}$, Int $k$) $\rightarrow$ Bool \;
\nl \Begin{
    \nl \lIf{$\left|\V(\mathcal{P})\right| + k > \left|\V(\mathcal{T})\right|$}{$\KwSty{return}~\KwSty{false}$\label{line:enough}}
    \nl $L \gets \big[ (\mathcal{P},\ \mathcal{T}), (\shortoverline{\mathcal{P}},
        \shortoverline{\mathcal{T}}),$
        \\\hspace*{1.0em}$(\mathcal{P}^{1 + k,2},\ \mathcal{T}^{1,2}), (\mathcal{P}^{2 + k,2},\
    \mathcal{T}^{2,2}), (\mathcal{P}^{3 + k,2},\ \mathcal{T}^{3,2}) \big]$ \;
    \nl \ForEach{$v \in \V(\mathcal{P})$}{
        \nl $D_v \gets \V(\mathcal{T})$\;
        \nl \ForEach{$(P,\,T) \in L$}{
            \nl $D_v \gets D_v \cap \{ w \in D_v : $ \\
            \hspace*{7em}$v \sim_P v \Rightarrow w \sim_T w~\wedge$ \\
            \hspace*{7em}$\nds_P(v) \lesspreceq{k} \nds_T(w)\}$ \;
        }
        \nl $D_v \gets D_v \cup k~\textnormal{distinct wildcard values}$
    }
    \nl \lIf{\FuncSty{propagate}(D)}{$\KwSty{return}~\FuncSty{search}(L, D, k)$}
    \nl \lElse{$\KwSty{return}~\KwSty{false}$}
}
\bigskip
\nl $\FuncSty{search}$ (GraphPairs $L$, Domains $D$, Int $k$) $\rightarrow$ Bool \;
\nl \Begin{
    \nl \lIf{$D = \emptyset$}{$\KwSty{return}~\KwSty{true}$}
    \nl $D_v \gets \textnormal{the smallest domain in}~D$ \;
    \nl \ForEach {$v' \in D_v~\textnormal{ordered by static degree in}~\mathcal{T}$\label{line:parforeach}}{
        \nl \If{$v'$ \textnormal{is not the first wildcard we have tried}}{
            \nl \KwSty{continue} \;
        }
        \nl $D' \gets \FuncSty{clone}(D)$ \;
        \nl $D'_v \gets \{ v' \} $ \;
        \nl \If{\FuncSty{propagate}(D')}{
            \nl \If{$\FuncSty{search}(L, \{ D \in D' : \left|D\right| > 1 \}, k)$}{
                \nl $\KwSty{return}~\KwSty{true}$ \;
            }
        }
    }
}
\bigskip
\nl $\FuncSty{propagate}$ (GraphPairs $L$, Domains $D$) $\rightarrow$ Bool \;
\nl \Begin{
    \nl \While{\KwSty{true}}{
        \nl \If{\textnormal{no unit domains remain}}{
            \nl \If{$\KwSty{not}~\FuncSty{allDifferent}(D)$}{
                \nl $\KwSty{return}~\KwSty{false}$ \;
            }
        }
        \nl \lIf{$\textnormal{no unit domains remain}$}{$\KwSty{return}~\KwSty{true}$}
        \nl $D_v = \{ v' \} \gets \textnormal{a unit domain from}~D$ \;
        \nl \ForEach{$D_w \in D \setminus \{ D_v \}$}{
            \nl $D_w \gets D_w \setminus \{ v' \}$ \label{line:removev} \;
            \nl \ForEach{$(P,\,T) \in L$}{
                \nl \If{$v \sim_P w$}{
                    \nl $D_w \gets D_w \cap \left(\N_T(v') \cup \textnormal{wildcards}\right)$
                }
            }
            \nl \lIf{$D_w = \emptyset$}{$\KwSty{return}~\KwSty{false}$}
        }
    }
}
\bigskip
\nl $\FuncSty{allDifferent}$ (Domains $D$) $\rightarrow$ Bool \;
\nl \Begin{
\nl $(F,\,H,\,A,\,n) \gets (\emptyset,\,\emptyset,\,\emptyset,\,0)$ \;
\nl \ForEach{$D_v \in D$ \textnormal{from smallest to largest\label{line:eachdomain}}}{
    \nl $F \gets F + v$ \;
    \nl $D_v \gets D_v \setminus H$ \label{line:elimhall} \;
    \nl $(A,\,n) \gets (A \cup D_v,\,n + 1)$ \label{line:acc} \;
    \nl \lIf{$D_v = \emptyset~\KwSty{or}~\left|A\right| < n$}{$\KwSty{return}~\KwSty{false}$\label{line:failhall}}
    \nl \lIf{$\left|A\right| = n$}{$(H,\,A,\,n) \gets (H \cup A,\,\emptyset,\,0)$\label{line:hall}}
}
\nl $\KwSty{return}~\KwSty{true}$ \;
}
\end{multicols}
\vspace*{0.4cm}% bug in algorithm2e + multicol: incorrect space left above the vertical line
\caption{An algorithm for the $k$-less subgraph isomorphism problem.}
\label{algorithm:sip}
\end{algorithm*}

>>>>

?? Also the clique encoding.

@@

?? Using the French code, and the Glasgow clique code.

Note that many of these instances are \emph{much} larger than were used in the recent comparison of
maximum common subgraph algorithms by \citet{DBLP:conf/cp/McCreeshNPS16}: for unlabelled graphs,
they used pairs of graphs with up to 35 vertices in each, whilst our dataset contains patterns with
up to ?? vertices and targets with up to ?? vertices. This causes problems: both the maximum common
subgraph algorithm and the clique encoding require $O(\left|\V(P)\right|^2\left|\V(T)\right|^2)$
memory. For maximum common subgraph, 1,560 of the instances cannot fit in the amount of RAM we have
available---we treat these instances as having timed out. The situation for the clique encoding is
even worse, and 3,653 of these instances do not fit in RAM. In contrast, the approach we discuss
here requires only $O(\left|\V(P)\right|^2\left|\V(T)\right|)$ space (which would still be a problem
on some desktop machines, but not on our server hardware).

@@

\begin{algorithm*}[tb]
\DontPrintSemicolon
\begin{multicols}{2}
\nl $\FuncSty{klessSubgraphIsomorphism}$ (\\\hspace*{2em}Graph $\mathcal{P}$, Graph $\mathcal{T}$, Int $k$) $\rightarrow$ Bool \;
\nl \Begin{
    \nl \lIf{$\left|\V(\mathcal{P})\right| + k > \left|\V(\mathcal{T})\right|$}{$\KwSty{return}~\KwSty{false}$\label{line:enough}}
    \nl $L \gets \big[ (\mathcal{P},\ \mathcal{T}), (\shortoverline{\mathcal{P}},
        \shortoverline{\mathcal{T}}),$
        \\\hspace*{1.0em}$(\mathcal{P}^{1 + k,2},\ \mathcal{T}^{1,2}), (\mathcal{P}^{2 + k,2},\
    \mathcal{T}^{2,2}), (\mathcal{P}^{3 + k,2},\ \mathcal{T}^{3,2}) \big]$ \;
    \nl \ForEach{$v \in \V(\mathcal{P})$}{
        \nl $D_v \gets \V(\mathcal{T})$\;
        \nl \ForEach{$(P,\,T) \in L$}{
            \nl $D_v \gets D_v \cap \{ w \in D_v : $ \\
            \hspace*{7em}$v \sim_P v \Rightarrow w \sim_T w~\wedge$ \\
            \hspace*{7em}$\nds_P(v) \lesspreceq{k} \nds_T(w)\}$ \;
        }
        \nl $D_v \gets D_v \cup k~\textnormal{distinct wildcard values}$
    }
    \nl \lIf{\FuncSty{propagate}(D)}{$\KwSty{return}~\FuncSty{search}(L, D, k)$}
    \nl \lElse{$\KwSty{return}~\KwSty{false}$}
}
\bigskip
\nl $\FuncSty{search}$ (GraphPairs $L$, Domains $D$, Int $k$) $\rightarrow$ Bool \;
\nl \Begin{
    \nl \lIf{$D = \emptyset$}{$\KwSty{return}~\KwSty{true}$}
    \nl $D_v \gets \textnormal{the smallest domain in}~D$ \;
    \nl \ForEach {$v' \in D_v~\textnormal{ordered by static degree in}~\mathcal{T}$\label{line:parforeach}}{
        \nl \If{$v'$ \textnormal{is not the first wildcard we have tried}}{
            \nl \KwSty{continue} \;
        }
        \nl $D' \gets \FuncSty{clone}(D)$ \;
        \nl $D'_v \gets \{ v' \} $ \;
        \nl \If{\FuncSty{propagate}(D')}{
            \nl \If{$\FuncSty{search}(L, \{ D \in D' : \left|D\right| > 1 \}, k)$}{
                \nl $\KwSty{return}~\KwSty{true}$ \;
            }
        }
    }
}
\bigskip
\nl $\FuncSty{propagate}$ (GraphPairs $L$, Domains $D$) $\rightarrow$ Bool \;
\nl \Begin{
    \nl \While{\KwSty{true}}{
        \nl \If{\textnormal{no effectively-unit (treating all wildcard values as a single value)
        domains remain}}{
            \nl \If{$\KwSty{not}~\FuncSty{allDifferent}(D)$}{
                \nl $\KwSty{return}~\KwSty{false}$ \;
            }
        }
        \nl \If{\textnormal{no effectively-unit domains remain}}{
            \nl $\KwSty{return}~\KwSty{true}$ \;
        }
        \nl $D_v \gets \textnormal{an effectively-unit domain from}~D$ \;
        \nl $v' \gets $ the single value in $D_v$, or an arbitrary wildcard value if there is more
        than one \;
        \nl \ForEach{$D_w \in D \setminus \{ D_v \}$}{
            \nl $D_w \gets D_w \setminus \{ v' \}$ \label{line:removev} \;
            \nl \ForEach{$(P,\,T) \in L$}{
                \nl \If{$v \sim_P w$}{
                    \nl $D_w \gets D_w \cap \left(\N_T(v') \cup \textnormal{wildcards}\right)$
                }
            }
            \nl \lIf{$D_w = \emptyset$}{$\KwSty{return}~\KwSty{false}$}
        }
    }
}
\bigskip
\nl $\FuncSty{allDifferent}$ (Domains $D$) $\rightarrow$ Bool \;
\nl \Begin{
\nl $(F,\,H,\,A,\,n) \gets (\emptyset,\,\emptyset,\,\emptyset,\,0)$ \;
\nl \ForEach{$D_v \in D$ \textnormal{from smallest to largest\label{line:eachdomain}}}{
    \nl $F \gets F + v$ \;
    \nl $D_v \gets D_v \setminus H$ \label{line:elimhall} \;
    \nl $(A,\,n) \gets (A \cup D_v,\,n + 1)$ \label{line:acc} \;
    \nl \lIf{$D_v = \emptyset~\KwSty{or}~\left|A\right| < n$}{$\KwSty{return}~\KwSty{false}$\label{line:failhall}}
    \nl \lIf{$\left|A\right| = n$}{$(H,\,A,\,n) \gets (H \cup A,\,\emptyset,\,0)$\label{line:hall}}
}
\nl $\KwSty{return}~\KwSty{true}$ \;
}
\end{multicols}
\vspace*{0.4cm}% bug in algorithm2e + multicol: incorrect space left above the vertical line
\caption{An algorithm for the $k$-less subgraph isomorphism problem.}
\label{algorithm:sip}
\end{algorithm*}
